---
title: "BANK CAMPAIGN ANALYSIS"
author: "Nolan Cardozo"
date: "25 January 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data

The `Bank Campaign` data set contains a campaign run by the bank and the response from the campaign. It contains a collection of variable given as follows.

Some of the variables in the `Bank Campaign` data set are:
***(Also classifying them into(Dimensions, Metrics, Dates, Text and Geo))***

* **Age Group** - The Age Group of the respondents of the campaign.**(Dimension)** 

* **Job** - The Job Title of the respondents of the campaign.**(Dimension)**

* **Marital** - The Marital Status of the respondents of the campaign.**(Dimension)**

* **Education** - The Highest Educational Qualification of the respondents of the campaign. **(Dimension)**
* **Balance Group** - The income group to which the respondents of the campaign belongs to. **(Dimension)**
* **Month** - The Month in which the campaign was run. **(Date)**

* **Response Rate** - It captures data whether the respondent responded to the campaign or not. **(Can be both Dimension and metric but makes more sense as dimension)**


##Reading the dataset into R
```{r,out.width=15}
Bank.Campaign <- read.csv("C:/Users/Admin/Desktop/data visulisation/Bank Campaign.csv")
```

##Inspecting the first and last 6 rows of the data for better undertanding of contents
```{r,out.width=20}
head(Bank.Campaign)
tail(Bank.Campaign)
```

##Inspecting the structure and frequency of the data

```{r}
str(Bank.Campaign)
```

##The number of rows and columns in the dataset

```{r}
dim(Bank.Campaign)
```

##Running basic summary statistics on the data 

```{r}
summary(Bank.Campaign,maxsum=15)
```
##Checking for NA's in the dataset visually using package VIM and function aggr
```{r, include=FALSE}
library(VIM)
```
```{r, echo=TRUE, fig.width=15}
missing_plot<-aggr(Bank.Campaign,col=c('blue','red'),
                      numbers=TRUE,sortVars=TRUE,
                      labels=names(Bank.Campaign),gap=1,
                      ylab=c("Missing Data","Pattern"), plot=TRUE)
```
#####The above head map and histogram shows no null values in the dataset

###The count of all non null values in each column
```{r}
as.data.frame(sapply(Bank.Campaign,function(y) sum(length(which(!is.na(y))))))
```
#####The above figure shows no null values in the dataset 

##The frequency plots for all the dimensions are as follows:

```{r, echo=FALSE}
library(ggplot2)
ggplot(data= Bank.Campaign, aes(Age.Group )) + geom_bar(fill= "steelblue") + labs(x="Age group",y= "Number of campaigns") + ggtitle("Number of campaigns per Age group")
```


```{r, echo=FALSE, fig.height=6, fig.width=10}
ggplot(data= Bank.Campaign, aes(Job)) + geom_bar(fill= "steelblue") + labs(x="Job",y= "Number of campaigns") + ggtitle("Number of campaigns per Job Type")
```


```{r, echo=FALSE}
ggplot(data= Bank.Campaign, aes(Marital)) + geom_bar(fill= "steelblue") + labs(x="Marital Status",y= "Number of campaigns") + ggtitle("Number of campaigns per Marital Status")
```


```{r, echo=FALSE}
ggplot(data= Bank.Campaign, aes(Education)) + geom_bar(fill= "steelblue") + labs(x="Education",y= "Number of campaigns") + ggtitle("Number of campaigns per Educational backround")
```


```{r, echo=FALSE}
ggplot(data= Bank.Campaign, aes(Default)) + geom_bar(fill= "steelblue") + labs(x="Default",y= "Number of campaigns") + ggtitle("Number of campaigns per Default rate")
```


```{r, echo=FALSE}
ggplot(data= Bank.Campaign, aes(Balance.Group)) + geom_bar(fill= "steelblue") + labs(x="Balance Group",y= "Number of campaigns") + ggtitle("Number of campaigns per Balance Group")
```


```{r, echo=FALSE}
ggplot(data= Bank.Campaign, aes(Month)) + geom_bar(fill= "steelblue") + labs(x="Month",y= "Number of campaigns") + ggtitle("Number of campaigns per Month")
Bank.Campaign$Response.Rate <-
factor(Bank.Campaign$Response.Rate)
```


```{r, echo=FALSE}
ggplot(data= Bank.Campaign, aes(Response.Rate)) + geom_bar(fill= "steelblue") + labs(x="Response Rate",y= "Number of campaigns") + ggtitle("Number of campaigns per Response Rate")
```

